---
title: "rnotes"
output: 
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
    font-size: 30  
  theme: !expr sample(c('cosmo','united','yeti'),1)
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R 

## **Introduction to Calculation**
### Calculator
```{r , eval=TRUE}
c(9+6,9*6,9/6,9^6,9%%6,9%/%6,abs(-9))
```
### Sequence
```{r , eval=TRUE}
1:10;seq(1,10,2)
```
### Exponential & logrithm
```{r eval=TRUE}
c(exp(10),log(2),log2(2),log10(10))
```
### Approximated values
```{r}
c(round(165.96253),round(165.96253,digit=3),round(165.96253,digit=-2))

c(signif(165.96253,digit=2),floor(165.96253),ceiling(176.96253),trunc(165.96253))
```
### Descriptive statistic
```{r}
x<-1:25
summary(x)
#If there is a missing value in the set,then the summary will all be NA.Therefore,we should set na.rm=TRUE.
```

## **Vectors**
### Numeric vector

 * Use __seq()__ and __rep()__ to construct a vector with similar patterns.
```{r}
seq(1,5,by=0.5)
seq(1,10,length=5)
rep(3,10)
rep(seq(2,10,2),2)
```

 * Use __unique()__ to return the elements that appear at least one time.

 * Use __table()__ to count the number of each element.
```{r}
b<-c(1,2,2,2,1,2,3,3,7,2,3,4,1,7)
unique(b)
table(b)
```

 * Use __union__(vector1,vector2) to find all elements between two vectors.

 * Use __intersect__(v1,v2) to find the same elements between two vectors.

 * Use __setdiff__(v1,v2) to find the difference between two vectors but only find the different elements in v1.

 * Use __setequal__(v1,v2) to find if v1 is the same as v2.
```{r}
B<-c(4,1,7,2,3);C<-c(12,2,3,1)
union(B,C)
intersect(B,C)
setdiff(B,C)
setequal(B,C)
```

 * Use __sort()__ or __order()__ to sort a vector or grt the position or sorting a vector.

```{r}
d<-c(3,4,2,5,7,23,4,8,11)
sort(d)
sort(d,decreasing=TRUE)
inc.order<-order(d)
dec.order<-order(d,decreasing=TRUE)
inc.order
d[inc.order]
dec.order
d[dec.order]
```

 * To __change values__ in a vector
```{r}
d<-c(3,4,2,5,7,23,4,8,11)
d[5]<-6
d[7]<-2
d
```

 * To __delete elements__ from a vector
```{r}
vt0<-c(1:5) ##create a vector
vt0[-3]
```

### Character vector

-The vector is made up of strings of characters.
-All elements of a vector must be of the same type.
```{r}
colors<-c('yellow','blue','RED','white')
more.colors<-c(colors,'pink','purple')
more.colors
```
 * Use __substr()__ to take substring.

 * Use __grep()__ to search a string from a vector

 * Use __paste()__ to build up strings.

```{r}
colors<-c('yellow','blue','RED','white')
more.colors<-c(colors,'pink','purple')
substr(colors,1,3)
grep('e',more.colors)
paste(colors,'line')  ##with space
paste(colors,'line',sep="")  ##with no space
 
```
__factor()__:factor(x=character())
```{r}
colors<-c('yellow','blue','RED','white')
fc<-factor(colors)
fc
levels(fc)
as.integer(fc) ##transform character into integer
```

### Logical vector

-Logical words:True(T) or False(F)

1.Boolean operations
-> Use __sum()__ to count how many TRUE in a vecotr.
```{r}
D<-c(T,T,T,F,F,T)
sum(D)
sum(!D) ##count how many FALSE
```

2.Type of __is. series__

 * is.null():check if it is a null vector

 * is.na():check if there is a missing value

 * is.nan():check if there is not a number in a vector

 * is.infinite():check if a vector is infinite

 * is.character():check if value in a vector is the type of character

 * is.integer():check if value in a vector is the type of integer

 * is.numeric():check if value in avector is the type of number

```{r}
x1<-1:5
x2<-c(1:3,2.2)
x3<-c('red','blue','white')
c(is.integer(x1),is.numeric(x1))
c(is.character(x3),is.null(x2))
```
3.Use "class()" to identify the type of a vector


## **Matrix**

 * Use __rbind()__ or __cbind()__ to construct
```{r}
a<-1:5;b<-6:10;c<-11:15;d<-16:20
M1<-rbind(a,b,c,d) ##by row
M1;dim(M1)
M2<-cbind(a,b,c,d) ##by column
M2;dim(M2)
```

 * Construct by a __long vector__
```{r}
X<-1:20
matrix(X,nrow=5,ncol=4,byrow=TRUE)
```

 * Use __column names__ or __row names__ to add names
```{r}
colnames(M1)<-c('V1','V2','V3','V4','V5')
dimnames(M1)
```

 * Use __[ ]__ to __extract elements__ from a matrix
```{r}
M1[2,4]
M1[,3]
M1[c(2,4), ]
M1[10]
M1[c('b','c'), c('V1','V5')]
```

 * Use __t()__ to create transpose matrix
 
 * Use **%*%** to evaluate multiplication of matrices
 
 * Use __diag()__ to make diagonal matrix
 
 * Use __det()__ to count determinant
 
 * Use __solve()__ to create inverse matrix
 
 * Use __eigen()__ to evaluate eigen value and eigen vector
```{r}
A <- matrix(c(1:4),2,2)
B <- matrix(c(5:8),2,2)
t(A)
A %*% B
diag(A)
det(B)
solve(B)
eigen(A)
```
 
## **Controlflow**

### For() loop

 * allows one to specify that a certain operation should be repeated a fixed number of times.

**syntax**

**`for(index in vector) commands` **

where

 * index: the index is changed each time
 * vector: it includes all possible cases
 * commands: actions 

```{r}
## evaluate n!
n<-10
ans <- 1
for (j in n:1){
  ans <- ans*j
}
ans
```

### if() statement

**syntax**

**`if(condition) {commands when TRUE}`**

**`if(condition) {commands when TRUE} else {commands when FALSE}`**

(The else part is __optional__)

**format** could be :

if(condition1){

commands when condition1 is TRUE

} else if (condition2){

commands when condition2 is TRUE

}else{

commands when condition1 and condition2 are FALSE


}

```{r}
basket <- c('r','b','y','y','y','r','r','r','r')
n.r <-0; n.b<-0; n.y <-0
for(i in 1:length(basket)){
  if(basket[i]=='r'){
    n.r <- n.r+1
  }else if(basket[i]=="b"){
    n.b <- n.b + 1
  }else{
    n.y <- n.y + 1
  }
}
c(n.r, n.b, n.y)
```

### While() loop

 * The pattern of repetition isn't known in advance and we need to do some calculations and keep going as a condition holds.

 * while(): need the initial, condition of stopping, and iteration or replacement

**syntax**

**`while(condition) statements`**

```{r}
## fibonacci number
fib1<-1
fib2<-1
fibonacci <- c(fib1)
while (fib2 < 100){
  fibonacci <- c(fibonacci, fib2)
  fib1 <- fibonacci[length(fibonacci) - 1]
  fib2 <- fib1 + fibonacci[length(fibonacci)]
}
fibonacci
```


### Break, repeat, and next

 * **break** : cause the loop to terminate immediately
 
**`if (condition) break`**

 * **repeat** : repeat until we execute a break statement
 
**`repeat {statements}`**

 * **next** : cause control to return immediately to the top pf the loop
 
```{r}
Fib1 <- 1
Fib2 <- 2
Fibonacci <- c(Fib1)
repeat
{
  if(Fib2 >= 100) break
    Fibonacci <- c(Fibonacci, Fib2)
    Fib1 <- Fibonacci[length(Fibonacci) - 1]
    Fib2 <- Fib1 + Fibonacci[length(Fibonacci)]
}
Fibonacci
```


## **Functions**

### Source()

 * After creating and **saving** the script,use it to run the code
 * Use __print()__ r __cat()__ to show the final answer

#### Different from __print()__ and __cat()__

 * **print()** shows with " " between characters
 * **cat()** shows without " " between characters
 
```{r}
k <-1
while(k<100){
  fac <-c()
  for(j in 1:k){
    if((k %% j) == 0){
      fac <- c(fac,j)
    }
  }
  if (k == sum(fac[-length(fac)])){
    cat(paste(k,"=", sep=" "),
        paste(fac[-length(fac)], collapse= "+"), "\n")
  }
  k<- k+1
}
```


### With one single output

```{r}
f <- function(x){
  4*x+17
}
f(3)
```

 * **return()** or **invisible()**

(if you don't put __return()__ at the end of the function,it will treat the last line as the result automatically.)

```{r}
make.percent <- function(x){
  x.percent <- round(100*x, digit=1)
  x.show <- paste(x.percent, "%", sep=" ")
  return(x.show)
}
make.percent(c(0.1222,0.0417,0.0221,0.0023))
```


### With multiple input variables

```{r}
perfect.num<- function(a,b){
  k <-a
  res <- c()
  while (k<-b){
    fac <-c()
    for(j in 1:k){
      if((k %% j) == 0) fac <-c(fac,j)
    }
    if (k == sum(fac[-length(fac)])){
      res <- c(res,k)}
      k <- k+1
  }
  return(res)
}
#perfect.num(1,10)
```

### Return multiple output

#### __list()__ : a collection of all type of objects

```{r}
g<- list(a=3,b=c(1,2,3,4,5),c=c("A","B"))
g
```

####  __Extract an element__ from a list

 * **$** : the name of the element
 * **[[ ]]** : the location or the name of the element

```{r}
g <- list(a=3 , b= c(1,2,3,4,5), c=c("A","B"))
g$a
g[[3]]
```

#### Use function() and list()

```{r}
perfect.num <- function(a,b){
  k<-a
  res<-c()
  ans <- list()
  while (k<b){
    fac <- c()
    for(j in 1:k){
      if ((k%%j) ==0) fac <-c(fac,j)
    }
    if (k == sum(fac[-length(fac)])){
      res <- c(res,k)
      ans[[paste(k)]] <- fac[-length(fac)]
    }
      k<- k+1
  }
  ans$num <- res
  return(ans)
}
ANS <- perfect.num(1,10)
ANS
```


## **DataFrame**

### Data.frame()

 * Use __data.frame()__ to construct data frames from vectors
 
 * Use __str()__ to examine the types of columns in the data frame
 
```{r}
colorss <- c("red","yellow","white")
numbers<-c(1,2,3)
colorss.and.numbers <- data.frame(colorss,numbers,more.numbers=c(4,5,6))
colorss.and.numbers

###the difference from the matrix
#x<-matrix(NA,3,3) 這行已經判定矩陣內的元素為文字形
#x[,1]<-c("red","yelloe"."white")
#x[,2]<-c(1,2,3)
#x[,3]<-c(4,5,6)
#colnames(x)<-c("colors","numbers","more.numbers")
#x

str(colorss.and.numbers)
```
 
 * Use __stringAsFactor = FLASE__ to change the type from factor to character

```{r}
colorss.and.numbers <- data.frame(colorss,numbers,more.numbers=c(4,5,6),stringsAsFactors = FALSE)
str(colorss.and.numbers)
```

 * Add column and row vectors into a dataframe
   * column : use _$_ or __cbind()__ (matrix can't use _$_)
   * row : use __rbind()__
   
```{r}
colorss <- c("red","yellow","white")
number1<-c(1,2,3)
number2 <- c(10,11,12)
colors.and.numbers <- data.frame(colorss,number1)
colors.and.numbers $ more.numbers <-c(4,5,6) ##add one column
colors.and.numbers <-cbind(colors.and.numbers,number2)
colors.and.numbers
```
   
### Data input and output at hard disk

 * Change the dirtectory: __setwd("dir")__
 * Find the directory: __get()__
 * **dump**(objects,file="xxx.R")
 * Use __load()__ to retrieve data
 * Use __write.table__(X,file="xxx.txtx")
 * Use __write.csv__(X,file="xxx,csv)

## **Graphics**

### Statistic Figures

 * Histogram: hist(x)
 * Boxplot: boxplot(x)
   -> 中間線為中位數,圖形上下短線為1.5IQR
 * Density plot: plot(density(x))
 * Pie chart: pie(x)
 * Barchart: barplot(x)
 * **plot(x,y)** : scatter plot for the correlation between teo variables(**cor(x,y)** : the correlation of two variables)
 * **pairs()** to see scatter plots for nultiple variables
 
### Create a plot region with specifical items

 * **Type**: "p":points, "l": lines, "b": both lines and points, "o": lines pass through points, "h": the lines between points and x-axis, "n": no content
 
 * **par(mfrow= c(2,3))** ## split plot ranges
 * **pch**: change the type of points
 * **cex**: size of points
 * **cex.main** : size of the title
 * **cex.axis** : size of the labels of ticks of an axis
 * **cex.lab** : size the labels of the name of an axis
 * **lty** : the style of lines
 * **lwd** : the width of lines
 * **col** : colors
 * **main** : main title(at the top)
 * **sub** : sub title(at the bottom)
 * **x,ylab** : name of x-axis,y-axis
 * **x,ylim** : ranges of variables
 * **curve(expression,range)**: the curve of function

### Add some objects to the existing plot

 * **points()** : add points
 
 * **lines()** : add lines
 
 * **abline()** : add straight lines
 
 * **layout()** : split the whole plot by some locations of plot

## **Apply**

### Function : apply()

#### Return a vector or array or list of values obtained by applying a finction to margins of __an array or a matrix__

**apply(X, MARGIN, FUN,...)**

 * **X**: array or matrix
 * **MARGIN**: for a matrix 1:rows, 2:columns
 * **FUN**: name of a function
```{r}
A <- matrix(rnorm(10),5,2)
apply(A,1,sum); rowSums(A)
```
 
### Function: lapply()

#### Returns a list of the same length as X, each element of which is the result of applying FUN to the corresponding element of X

**lapply(X,FUN...)**

 * **X**: vector or data frame of list
 * **FUN**: name of a function
```{r}
cc<- list(a=1:10,beta=exp(-3:3))
lapply(cc,mean)
```

### Function: sapply()

#### Similarly to lassply(X,FUN,...) if sapply(x,f,simplify=FALSE, USE.NAMES=FALSE)

**sapply(X,FUN,...simplify=TRUE)**

 * **X**: vector or data frame or list
 * **FUN** : name of a function
 * **simplify** : logical; TRUE will return an array or a matrix with the same length of output
 
```{r}
sapply(cc,summary,simplify = FALSE)
```
 
```{r}
d<-list(a=sample(1:10,10,replace=TRUE),
        b=sample(1:5,20,replace=TRUE))
sapply(D,unique)
sapply(D,table)

```

